---
title: "Shiny App"
output: html_document
https://www.kaggle.com/dorbicycle/world-foodfeed-production/home
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# load libraries
```{r libraries}
library(tidyr)
library(devtools)
library(DataExplorer)
library(rvest)
library(dplyr)
library(lubridate)
library(stringr)
library(readr)
library(timeDate)
library(shiny)
library(plotly)
library(ggplot2)
library(ggthemes)
library(Hmisc)
library(microbenchmark)
library(knitr)
library(shinythemes)
library(shinyWidgets)
library(rgdal)
library(maps)
library(rworldmap)
library(DT)
library(scales)
```

#Import Data
```{r import data}
# UN food production data
data <- read_csv('C:/Users/BieniekJon/Desktop/901/R/Shiny/FAO.csv')
print(data)

# UN countries by region
region <- read_csv('C:/Users/BieniekJon/Desktop/901/R/Shiny/Region list.csv')
print(region)

# Food categories
categories <- read_csv('C:/Users/BieniekJon/Desktop/901/R/Shiny/Food Categories.csv')
print(categories)
```

# Use left join to add region and food categories to the UN data set
```{r join data}
joined_data <- left_join(data,region,by = c("Area" = "Country"))
print(joined_data)

joined_data2 <- left_join(joined_data,categories,by = c("Item" = "Items"))
print(joined_data2)
```

Remove and row with an NA
```{r remove NAs}
no_na <- joined_data2[complete.cases(joined_data2),]
no_na[is.na(no_na$`Food_Sub_Category`),]
```

Subset the data
```{r}
columns_of_interest <- c("Element","Area","Region","Food_Sub_Category", 
                         "Y1961", "Y1962", "Y1963", "Y1964", "Y1965","Y1966",
                         "Y1967","Y1968","Y1969","Y1970","Y1971","Y1972","Y1973",
                         "Y1974","Y1975","Y1976","Y1977","Y1978","Y1979","Y1980",
                         "Y1981","Y1982","Y1983","Y1984","Y1985","Y1986","Y1987",
                         "Y1988","Y1989","Y1990","Y1991","Y1992","Y1993","Y1994",
                         "Y1995","Y1996","Y1997","Y1998","Y1999","Y2000","Y2001",
                         "Y2002","Y2003","Y2004","Y2005","Y2006","Y2007","Y2008",
                         "Y2009","Y2010","Y2011","Y2012","Y2013")

data_subset <- no_na[columns_of_interest]
print(data_subset)
```

# Create a flat file using the gather() function.
```{r flat_file}
flat_all_yrs <- gather(data_subset, 
                              "Year", "Prod_one_tho_ton", 5:57)
print(flat_all_yrs)
```

Remove the letter Y from Year elements
```{r remove letter from string}
flat_all_yrs$Year <- gsub("Y", "",flat_all_yrs$Year)
print(flat_all_yrs)
```

Subset the years so the visualizations are more manageable
```{r subset years}
flat_file <- flat_all_yrs[flat_all_yrs$Year>=1993,]
print(flat_file)
```

Transform all columns except the Production column into factors. This will make ploting graphs easier.
```{r turn to factors}
flat_file$Element <- as.factor(flat_file$Element)
flat_file$Area <- as.factor(flat_file$Area)
flat_file$Region <- as.factor(flat_file$Region)
flat_file$Food_Sub_Category <- as.factor(flat_file$Food_Sub_Category)
flat_file$Year <- as.factor(flat_file$Year)
```

Shiny App
Variables for shiny
```{r shiny variables}
regions <- sort(unique(flat_file$Region))
food_cat <- sort(unique(flat_file$Food_Sub_Category))
```

Shiny code
```{r shiny app}
# Define UI for application that plots features of movies
ui <- shinyUI( 
    navbarPage("Global Food and Feed Production",   
               
               setBackgroundColor(
    color = c("white", "silver"),
    gradient = "radial",
    direction = c("top", "left")
                                  ),
               
    tabPanel("Insights",
                        
            fluidRow(
  
  # Sidebar layout with a input and output definitions
    # Inputs
    column(4,
           div(style = "padding: 20px 100px",
        pickerInput(inputId = "purpose", 
                     label = "Intended use of Goods Produced:", 
                     choices =c("Human Consumption" = "Food",
                                "Animal Consumption" = "Feed"),
                     options = list(`actions-box` = TRUE),
                     multiple = TRUE))
    ),
    
    column(4,
           div(style = "padding: 20px 100px",
        pickerInput(inputId = "region",
                    label = "Global Regions:",
                    choices = regions,
                    options = list(`actions-box` = TRUE),
                    multiple = TRUE))
    ),
    
    column(4,
           div(style = "padding: 20px 100px",
        pickerInput(inputId = "group",
                    label = "Food/Feed Category:",
                    choices = food_cat,
                    options = list(`actions-box` = TRUE),
                    multiple = TRUE))
    ),            
        
      # Outputs
        tabsetPanel(
          tabPanel("Stacked by Region",
                   plotlyOutput(outputId = "barplot",width = "100%", height = "500px")),
          tabPanel("Stacked by Food Category",
                   plotlyOutput(outputId = "barplot2",width = "100%", height = "500px")),
          tabPanel("Data Table",
                  DT::dataTableOutput("table")
                    )
                  )
                )
               ),

  
# Will be a Read Me
                tabPanel("README",
                         htmlOutput("text1")
                         )
              )
)
            
# Define server function required to create the scatterplot
server <- function(input, output,session) {
  
### reactive filtered data from inputs 
      filtered_data <- reactive({filter(flat_file, Element %in% input$purpose)})
      observe({
                updateSelectInput(session,"region", 
                                  choices = unique(filtered_data()$Region))
        })
      
      filtered_data_2 <- reactive({filter(filtered_data(), Region %in% input$region)})
      observe({
                updateSelectInput(session,"group", 
                                  choices = unique(filtered_data_2()$Food_Sub_Category))
      })
      
      filtered_data_3 <- reactive({filter(filtered_data_2(),Food_Sub_Category %in% input$group)})

### summarized data for first plot            
      plot_data1 <- reactive({filtered_data_3() %>% 
                                      group_by(Year,Region) %>% 
                                      summarise(sum = sum(Prod_one_tho_ton))})
      
### summarized data for second plot          
      plot_data2 <- reactive({filtered_data_3() %>% 
                                      group_by(Year,Food_Sub_Category) %>% 
                                      summarise(sum = sum(Prod_one_tho_ton))})

### Tab 1 plot 
      output$barplot <- renderPlotly(
     {
          p<-ggplot(plot_data1(), 
                    aes_string(x=plot_data1()$Year, 
                               y=plot_data1()$sum, 
                               fill=(plot_data1()$Region))) +
            theme_minimal() +
            geom_bar(stat="identity") +
            scale_y_continuous(labels = comma) +
            labs(title = "Global Production", 
                 subtitle = "Segmented by Region",
                 fill="",
                 x = "Years", 
                 y = "Tons(one thousands)" )
          ggplotly(p)   
   }
   )
      
### Tab 2 plot      
      output$barplot2 <- renderPlotly(
     {
          r<-ggplot(plot_data2(), 
                    aes_string(x=plot_data2()$Year, 
                               y=plot_data2()$sum, 
                               fill=(plot_data2()$Food_Sub_Category))) +
            theme_minimal() +
            geom_bar(stat="identity") +
            scale_y_continuous(labels = comma) +
            labs(title = "Global Production", 
                 subtitle = "Segmented by Food Category",
                 fill="",
                 x = "Years", 
                 y = "Tons(one thousands)" )
          ggplotly(r)   
   }
   )

### Table data - rename columns
table_data <- reactive({filtered_data_3()})
            
### Table
 output$table <- DT::renderDataTable({
    DT::datatable(table_data(), options = list(lengthMenu = c(10, 30, 50), pageLength = 10))
  })
    
### README tab
output$text1 <- renderUI(

    {

      HTML("<h3>About this Application</h3>
      <p>This application consists of data from the United Nation's Food and Agriculture Organiation ('FAO'). A link to the FAO websit can be found at the bottome of this              page.</p>

      <p>According to its website, the FAO is a specialized agency that leads international efforts to defeat hunger.The FAO currently has over 194 member states 
        and works in over 130 countries. Many of these member states are featured in this app. For a complete list of countries included in this app please 
        see the data table within the insights section.</p>

      <p>This app was created to provide easy access to food and feed production data for over 174 countries. Food is defined as goods intended to be consumed by humans, While         feed is defined as goods intended to be consumed by animals.</p>

      <p>The data included in this app covers the time period between 1993 and 2013.</p>

      <p> To use the app simply navigate to the insights section and select the desired inputs. Once the inputs are selected two bar charts will be generated along with a table         of the charts underlying data. from the drop down lists within the insights section of the app.</p>

      <p>Please enjoy this app!</p>")

      }

    )

}

# Create the Shiny app object
shinyApp(ui, server)
```
```{r}
colnames(flat_file)
```

